---
title: "R tutorial for Life Science Mathematics"
author: "Chen-Yi Tu"
date: "September 29, 2015"
output: html_document
---

# Basic operation
```{r, echo=TRUE}
#Getting help
?read.csv
#or
help(read.csv)
```

Check example on a help topic
```{r}
example(mean)
```

Set working directory
```{r, eval=FALSE,echo=TRUE}
setwd("OneDrive/in_progress/course/LifeScienceMathematics/Rclass/")
library("deSolve")
```

# Part 1: File I/O
```{r, echo=TRUE}
# read data as data frame
pollock = read.csv('pollockGOA.csv',header=TRUE)
mode(pollock)
head(pollock)
```

Call the element in the data frame by column name
```{r}
pollock$Year
```

Assign the element in the data frame to different variable
```{r, eval=TRUE,echo=TRUE}
age1 = pollock$X1
year = pollock$Year
```
```{r, eval=TRUE,echo=TRUE}
age1
```

Convert the data frame into matrix
```{r,eval=TRUE,echo=TRUE}
pollock_mat = data.matrix(pollock)
mode(pollock_mat)
```

Check the matrix by the index
```{r,eval=TRUE,echo=TRUE}
year = pollock_mat[,1]
data = pollock_mat[,c(2:16)]
```

```{r,eval=TRUE,echo=TRUE}
year
```

```{r,eval=TRUE,echo=TRUE}
data
```

Convert matrix to data frame
```{r,eval=TRUE,echo=TRUE}
output = as.data.frame(data)
```

Save the data frame as another file
```{r,eval=TRUE,echo=TRUE}
write.table(output,file="pollock_matrix.csv",row.names=FALSE,col.name=FALSE)
```


# Part II: Graphic
```{r,eval=TRUE,echo=TRUE}
plot(data[,1],data[,6],xlab="Age1",ylab="Age6")
```

```{r,eval=TRUE,echo=TRUE}
barplot(data[,1],xlab="Age",ylab="million ind.") # barplot
```

```{r,eval=TRUE,echo=TRUE}
plot(year,data[,6],type="l",ylab="million ind.") # line plot
```

# Part III: data creation, type and structure
```{r,eval=TRUE,echo=TRUE}
# create a vector
a <- c(1,2,5.3,6,-2,4) # numeric vector
b <- c("one","two","three") # character vector
c <- c(TRUE,TRUE,TRUE,FALSE,TRUE,FALSE) # logical vector
```

```{r,eval=TRUE,echo=TRUE}
a
```

Create a sequence with setted interval
```{r,eval=TRUE,echo=TRUE}
times <- seq(0, 100, by=0.01) #create a sequence of data with interval=0.1
```

Create a matrix
```{r,eval=TRUE,echo=TRUE}
mat = matrix(c(1:9),3,3)
```

# Part IV: matrix and basic mathematics
Matrix operation
```{r,eval=TRUE,echo=TRUE}
matT = t(mat) # transpose
matDiag = diag(mat) # diagnoal
mat2 = mat %*% mat # multiplication
lambda = eigen(mat) # eigen value analysis
```

Basic mathematics and statistics 
```{r,eval=TRUE,echo=TRUE}
total = sum(mat)
subtotal =  colSums(mat)
log(mat)
exp(mat)
sin(pi/2)
```

Example: use deSolve to solve ODE
```{r,eval=FALSE,echo=TRUE}
t = times
state = c(X=1,Y=1,Z=1)
parameters = c(a=?,b=?,c=?)
```

# Part V: function
```{r,eval=FALSE,echo=TRUE}
Lorenz<-function(t, state, parameters) {
   with(as.list(c(state, parameters)),{
    # rate of change
    dX <- a*X + Y*Z
    dY <- b * (Y-Z)
    dZ <- -X*Y + c*Y - Z
    # return the rate of change
    list(c(dX, dY, dZ))
    }) 
}
```

# Part VI: Programming control structure
For loop
```{r,eval=TRUE,echo=TRUE}
newMat = matrix(0,3,3)

for (i in 1:nrow(mat))
  for (j in 1:ncol(mat)){

    newMat[i,j] = mat[i,j]+2
}
```

```{r,eval=TRUE,echo=TRUE}
newMat
```

If loop
```{r,eval=TRUE,echo=TRUE}
if (sum(newMat)>50) r=1 else r=0
```

```{r,eval=TRUE,echo=TRUE}
newMat
```
