---
title       : Data analysis with R
subtitle    : 
author      : Crystal Tu
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---
```{r setup, include=FALSE, cache=FALSE}
suppressPackageStartupMessages({
  library(knitr)
  library(RColorBrewer)
  library(dplyr)
})
opts_chunk$set(echo = FALSE, cache = FALSE)
fit100 <- function(fname) sprintf('<img src="assets/img/%s" class="fit100" />', fname)
fit50 <- function(fname) sprintf('<img src="assets/img/%s" class="fit50" />', fname)
fig <- function(fname, size) sprintf('<img src="assets/img/%s" style="max-width:%d%%;max-height:%d%%;" />', fname, size, size)
```

## Useful link

- Today's slide http://crystaltu.github.io/Ranalysis/index.html
- Morzilla Firefox https://mozilla.com.tw/
- DSP data center http://ckan.nispc.tw/organization/dsp-data-camp-at-nccu
- LibreOffice https://zh-tw.libreoffice.org/

---
## R basic Review 

- Basic command
```{r echo=TRUE,eval=FALSE}
getwd() # where we are now
```
```{r echo=TRUE,eval=FALSE}
setwd("path") # set working directory
```
```{r echo=TRUE,eval=FALSE}
help() # get help
example() # get example
```

---
## Data

<left>`r fig("295px-Iris_virginica.jpg",75)`</left>

- Anderson's Iris data
- http://vincentarelbundock.github.io/Rdatasets/csv/datasets/iris.csv
- 臺北市實價登錄租賃資料 & YouBike 微笑單車 & YouBike X Weather Data
- http://ckan.nispc.tw/organization/dsp-data-camp-at-nccu

---
## Read data
```{r echo=TRUE,eval=TRUE}
iris <- read.csv('iris.csv') # read iris data
str(iris)
```

---
## Computation

- +, -, *, /
- 基本邏輯運算符號 (>, >=, <, <=, ==, not !=, and &, or |)

---

## Labeling categorical variables

```{r echo=TRUE,eval=TRUE}
?factor
species <- factor(iris$Species, levels = c("setosa","versicolor","virginica"), 
                  labels = c("set","ver","vir"))
sample(species, size=10)
```

---

## Subseting

```{r echo=TRUE,eval=TRUE}
library(dplyr)
virginica <- filter(iris, Species == "virginica")  #只選 I. virginica
```
```{r echo=TRUE,eval=TRUE}
head(virginica)
```

---

## Pivot table with `dplyr`
```{r, echo=TRUE, eval=TRUE}
summarise(group_by(iris,Species), mean(Sepal.Length), mean(Sepal.Width), 
          mean(Petal.Length), mean(Petal.Width))
```

---

## Descriptive statistics

```{r echo=TRUE,eval=TRUE}
summary(iris)
```

--- &twocol
## Visualization

*** =left
```{r echo=TRUE,eval=FALSE}
pairs(iris[1:4], main = "Iris Data", 
      pch = 21, bg = c("red","green3","blue"))
```

*** =right
```{r echo=FALSE,eval=TRUE}
pairs(iris[1:4], main = "Iris data", pch = 21, 
      bg = c("red","green3","blue"))
```

---

## Save file for further use

```{r echo=TRUE,eval=FALSE}
write.csv(x)
```

--- .segue .dark

## Scenario

---

`r fig("nccu_oversupply.png",100)`

---

## Questions

- 租金和坪數間，有什麼關係？（每坪租金大概多少錢？）
- 如果有關，這個關係有多強？
- 房屋類型（公寓、華廈）、屋齡、隔間數量或主要用途（住家、住商用）有沒有影響？
- 要如何估計每一個因素的影響程度？
- 如何利用已知關係，找出適合自己的選擇？

--- .segue .dark

## Model conception and definition

--- .largecontent

## I. Simple regression model

- $$Y=\beta_0 +\beta_1 X +\epsilon$$
- X: predictor
- Y: response

```{r echo=TRUE,eval=FALSE}
lm(formula)
```

--- .largecontent

## Parameter estimation
- Least square method

<center>`r fig("Linear_least_squares_example2.svg",100)`</center>

--- .largecontent

## Goodness-of-fit

- R-square

$$R^2 = 1 - \frac{SS_{residual}}{SS_{total}} $$

---

## Example

```{r echo=TRUE,eval=FALSE}
#taipei_rent <- read.table('TaipeiRealPriceRent_bi5.csv',header=TRUE,sep=",") # for Windows
taipei_rent <- read.table('TaipeiRealPriceRent_utf8.csv',header=TRUE,sep=",")
oldtype <- unique(taipei_rent$建物型態)
type <- factor(taipei_rent$建物型態, levels=oldtype, labels=c("住宅","商辦","華廈","公寓","店面","套房","廠辦","其他","透天厝","倉庫","工廠"))
taipei_rent$建物型態 <- type

library(dplyr)
wenshan <- filter(taipei_rent, 鄉鎮市區 == "文山區",主要用途 == "住家用"); #選文山區建物資料
tab1 <- select(wenshan,建物移轉總面積.平方公尺.,總價.元.)
tab2 <- select(wenshan,交易標的,建物型態,建物移轉總面積.平方公尺.,總價.元.,有無附傢俱,有無管理組織,contains("建物現況格局"));
names(tab1) <- c("area","price")
names(tab2) <- c("target","type","area","price","furniture","management","room","hall","toilet","seperation")
```

```{r echo=FALSE,eval=TRUE}
#taipei_rent <- read.table('TaipeiRealPriceRent_bi5.csv',header=TRUE,sep=",") # for Windows
taipei_rent <- read.table('TaipeiRealPriceRent_utf8.csv',header=TRUE,sep=",")
oldtype <- unique(taipei_rent$建物型態)
type <- factor(taipei_rent$建物型態, levels=oldtype, labels=c("住宅","商辦","華廈","公寓","店面","套房","廠辦","其他","透天厝","倉庫","工廠"))
taipei_rent$建物型態 <- type

library(dplyr)
wenshan <- filter(taipei_rent, 鄉鎮市區 == "文山區",主要用途 == "住家用"); #選文山區建物資料
tab1 <- select(wenshan,建物移轉總面積.平方公尺.,總價.元.)
tab2 <- select(wenshan,交易標的,建物型態,建物移轉總面積.平方公尺.,總價.元.,有無附傢俱,有無管理組織,contains("建物現況格局"));
names(tab1) <- c("area","price")
names(tab2) <- c("target","type","area","price","furniture","management","room","hall","toilet","seperation")
```

--- &vcenter
## Simple linear regresion
```{r echo=TRUE,eval=TRUE}
# simple linear regression with price~area
model1 <- lm(price~area,data=tab1)
plot(tab1$price~tab1$area,xlab="Price",ylab="Area");abline(model1, col="red")
```

---

```{r echo=TRUE,eval=TRUE}
summary(model1) # R-square=0.4686 summary of model result
```

---
## Prediction

```{r echo=TRUE,eval=TRUE}
#make a new data set and predict the outcome
new <- data.frame(area=seq(50,100,10))
predict.lm(model1,newdata=new)
```

---

## II. Multi-linear regression

```{r echo=TRUE,eval=FALSE}
lm(Y~X1+X2+..+Xn)
```

- What is the relationship between Y and Xs?
- How many X need?
- How to estimate the goodness-of-fit?
- How accurate can the model predict?

---
## Qualitative predictor (factor() with levels)

```{r echo=TRUE,eval=FALSE}
library(ggplot2)
qplot(area,price,data=tab2) + facet_grid(furniture~seperation) + 
  geom_smooth(method="lm") + theme(text=element_text(family="STHeiti",size=14))
```
<center>`r fig("furniture-seperation.png", 50)`</center>

---
## Multiple regression with categorical variable
```{r echo=TRUE,eval=TRUE}
model2 <- lm(price~area+type+furniture+seperation,data=tab2)
summary(model2) #R-square=0.5050
```

---
## Predictions
```{r echo=TRUE,eval=TRUE}
forecast <- data.frame(area=50,type="住宅",furniture="有",seperation="有")
predict(model2,forecast)
```

---
## Missing data 

```{r echo=TRUE,eval=FALSE}
lm(formula,na.action=?)
```

---.segue .dark

## Remark

---
## Nonlinear relationship

<center>`r fig("1280px-Anscombe's_quartet_3.png",75)`</center>

---
## Interaction terms
[Simpson's paradox](http://vudlab.com/simpsons/)

---
## Correlation != Causation

<center>`r fig("greek_debt_FB.png", 100)`</center>

[Source](http://www.businessweek.com/magazine/correlation-or-causation-12012011-gfx.html)

---
## Summary
- Linear model: lm(), abline()
- Model forecast: predict()

---

## Resource for further study
- Quick-R http://www.statmethods.net/
- R-blogger http://www.r-bloggers.com/
- stackoverflow https://stackoverflow.com/tags/R
- R-help mailing list https://stat.ethz.ch/mailman/listinfo/r-help

--- .segue .dark

## Q & A

--- .segue .dark

## IT'S YOUR TURN AGAIN

---
## Scenario

<center>`r fig("youbike_raining.png",75)`</center>

---
## YouBike x Weather 

```{r echo=FALSE,eval=TRUE}
youbikeXweather <- read.table('ubike-weather-utf8.csv',header=TRUE,sep=",")
names(youbikeXweather)
```

---
## Question
- YouBike平均車輛數（或其他資料）會受天氣影響嗎？


